Sparse Function-space Representation \\ of Neural Networks

### 2023-05-14 paper/main.tex
made rl fig bigger
updated rl section

\subsection{Reinforcement learning}
We now detail how our method can be used in model-based RL.
We consider stochastic environments with states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \),
actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\) and transition dynamics
\(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
$\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$, where  $\noise_{t}$
is i.i.d. transition noise.
We consider the episodic setting where the system is reset to an initial state $\state_{0}$ at each episode and we
assume that there is a known reward function $r : \stateDomain \times \actionDomain \rightarrow \R$.
% Following the Markov decision process formulation \cite{bellmanMarkovianDecisionProcess1957a}, we
% denote the states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \) and actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\)
% of the sytem, the reward function $r : \stateDomain \times \actionDomain \rightarrow \R$, and transition dynamics
% \(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
% $\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$ where  $\noise_{t}$
% is i.i.d. transition noise.
The goal of RL is to find the policy \(\pi : \stateDomain \rightarrow \actionDomain\) that maximises the sum of discounted rewards
in expecation over the transition noise,
\begin{align} \label{eq-model-free-objective}
\policy^{*} = \arg \max_{\policy \in \policyDomain} J(\transitionFn, \policy) = \arg \max_{\policy \in \policyDomain} \mathbb{E}_{\noise_{0:\infty}} \left[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \right]
\quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t},
\end{align}
where $\gamma \in [0, 1]$ is a discount factor and $\Pi$ is a set of policies.

\textbf{Model-learning}
In model-based RL, the state transition data set \(\mathcal{D} = \{\{(s_{t},a_{t}), s_{t+1}\}^{T_{i}}_{t=1}\}_{i=0}^{N}\) is first used to
learn a dynamic model \(f_{\mathbf{w}})\) using supervised learning.
\cref{alg-mbrl} shows the typical model-based RL loop.

\begin{align} \label{eq-fast-update-mpc}
  \policy_{i+1}^{\text{PS}}(\state) &= \arg \max_{\action_{0}} \max_{\action_{1:\Horizon}}
J^{\Horizon}(\action_{0:\Horizon}, \transitionFn) + \stateValueFn(\state_{\Horizon+1})
\quad
  \stateValueFn(\state) &= \mathbb{E} \left[ \sum_{t=0}^{\infty}     \discount^{t} \rewardFn(\state_{t},\action_{t}) \mid \state_{0}=\state \right] %\label{eq-fast-update-mpc-old}
\end{align}

In model-based RL, we obtain the posterior over the dynamics \(p(f\mid\mathcal{D})\) after performing (approximate) Bayesian
inference given a state transition data set \(\mathcal{D} = \{\{(s_{t},a_{t}), s_{t+1}\}^{T_{i}}_{t=1}\}_{i=0}^{N}\).
\cref{alg-mbrl} shows the typical model-based RL loop.
Importantly, the dynamics are usually only updated after an episode $i$.

\textbf{Posterior sampling}
Given the posterior dynamics \(p(\transitionFn \mid \mathcal{D})\),
an effective and  common approach is to simply take the expecation over both the aleatoric and epistemic uncertainty,

\cite{osbandWhyPosteriorSampling2017,osbandMoreEfficientReinforcement2013}
\begin{align} \label{eq-posterior-sampling}
\policy_{i+1}^{\text{PS}} = \arg \max_{\policy \in \policyDomain} \left[ J(\transitionFn, \policy) \right] \quad \text{s.t. } \transitionFn \sim p(\transitionFn \mid \dataset_{0:i})
\end{align}

@@

 \end{tikzpicture}}
 \end{subfigure}
 \hfill
  \setlength{\figurewidth}{.28\textwidth}
  \setlength{\figureheight}{\figurewidth}
 \begin{subfigure}{.78\textwidth}
   % \tikz\node[minimum width=\textwidth,minimum height=.35\textwidth,fill=black!10,rounded corners=2pt]{RL trjectories or something interesting?};
   % \includegraphics[width=0.5\textwidth]{./fig/rl.pdf}
   \input{./fig/rl.tex}
 \end{subfigure}
 \hfill
 \caption{\textbf{Reinforcement learning experiments} Training curves showing that our model's uncertainty estimates improve sample efficiency in RL.
   Our method (blue/green) converges to the optimal solution in fewer environment steps than the baseline model-based RL method (magenta) and (DDPG) the model-free baseline  (yellow).}
 \label{fig:rl}
\end{figure}

>>>>

\subsection{Reinforcement learning}
A key challenge in RL is to balance the trade-off between exploration and exploitation.
That is, should the agent select an action that it knows will lead to a high reward (exploitation), or should it select an action that it has
not selected before, which could lead to either a higher or lower reward (exploration).
In this section, we detail how our method's principled uncertainty estimates can be used to balance this trade-off using a model-based RL algorithim.

We consider stochastic environments with states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \),
actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\) and transition dynamics
\(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
$\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$, where  $\noise_{t}$
is i.i.d. transition noise.
We consider the episodic setting where the system is reset to an initial state $\state_{0}$ at each episode and we
assume that there is a known reward function $r : \stateDomain \times \actionDomain \rightarrow \R$.
% Following the Markov decision process formulation \cite{bellmanMarkovianDecisionProcess1957a}, we
% denote the states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \) and actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\)
% of the sytem, the reward function $r : \stateDomain \times \actionDomain \rightarrow \R$, and transition dynamics
% \(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
% $\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$ where  $\noise_{t}$
% is i.i.d. transition noise.
% $\mathbb{E}_{\noise_{0:\infty}} \big[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \big]$
% \begin{align} \label{eq-model-free-objective}
% \policy^{*} = \arg \max_{\policy \in \policyDomain} J(\transitionFn, \policy) = \arg \max_{\policy \in \policyDomain} \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
% \quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t},
% \end{align}
The goal of RL is to find the policy \(\pi : \stateDomain \rightarrow \actionDomain\)
(from a set of policies $\Pi$) that maximises the sum of discounted rewards
in expecation over the transition noise,
\begin{align} \label{eq-model-free-objective}
J(\transitionFn, \policy) = \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
\quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}
\end{align}
where $\gamma \in [0, 1]$ is a discount factor.
In this work, we consider model-based RL where a model of the transition dynamics \(f_{\mathbf{w}} \approx \transitionFn\) is learned and then used by a planning algorithm.
\begin{align} \label{eq-fast-update-mpc}
  \policy^{\text{Greedy}}(\state) &= \arg \max_{\pi \in \Pi} J(f_{\mathbf{w}^{*}}, \pi) \quad   \\
  \policy^{\text{PS}}(\state) &= \arg \max_{\pi \in \Pi} J(\hat{f}, \pi)
\quad \text{s.t. } \hat{\transitionFn} \sim p(\transitionFn_{\mathbf{w}} \mid \dataset),
\end{align}
Our exact set up uses
Model Predictive Path Integral (MPPI) control
\cite{panSample2015}
\cite{williamsModel2017}
A principled approach
% Given the posterior dynamics \(p(\transitionFn \mid \mathcal{D})\),
% an effective and  common approach is to simply take the expecation over both the aleatoric and epistemic uncertainty,
\cite{osbandWhyPosteriorSampling2017,osbandMoreEfficientReinforcement2013}

\textbf{Appendix}

% \cref{alg-mbrl} shows the typical model-based RL loop.
\begin{align} \label{eq-fast-update-mpc}
  \policy_{i+1}^{\text{PS}}(\state) = \arg \max_{\action_{0}} \max_{\action_{1:\Horizon}}
\E \bigg[ \sum_{t=0}^{H-1} \gamma^{t} r(\state_{t},\action_{t}) \mid \state_{0}=\state  \bigg] + Q_{\theta}(\state_{\Horizon}, \action_{H})
\quad \text{s.t. } \state_{t+1} &= \hat{\transitionFn}(\state_{t}, \action_{t}) + \noise_{t} \\
\hat{\transitionFn} &\sim p(\transitionFn \mid \dataset),
\end{align}
with $\hat{\transitionFn} \sim p(\transitionFn \mid \dataset)$

We use deep deterministic policy gradient (DDPG) \cite{lillicrapContinuousControlDeep2016} to learn an action value function $Q_{\theta}$.
Note that we also learn a policy but its sole purpose is for learning the value function.

@@

 \end{tikzpicture}}
 \end{subfigure}
 \hfill
  \setlength{\figurewidth}{.4\textwidth}
  \setlength{\figureheight}{\figurewidth}
 \begin{subfigure}{.78\textwidth}
   % \tikz\node[minimum width=\textwidth,minimum height=.35\textwidth,fill=black!10,rounded corners=2pt]{RL trjectories or something interesting?};
   % \includegraphics[width=0.5\textwidth]{./fig/rl.pdf}
   \input{./fig/rl.tex}
 \end{subfigure}
 \hfill
 \caption{\textbf{Reinforcement learning experiments} Training curves showing that our model's uncertainty estimates improve sample efficiency in RL.
   Our method (blue/green) converges to the optimal solution in fewer environment steps than the baseline model-based RL method (magenta) and (DDPG) the model-free baseline  (yellow).}
 \label{fig:rl}
\end{figure}
